---
title: "Data Visualization in R"
output: html_notebook
---

# Set Working Directory

```{r}
setwd("/Users/nick/Documents/Teaching/workshops/data_science_series/4b_viz_r")
```


# Import the Data

```{r}
mpg <- read.csv("mpg.csv")
```

# One Variable Plots

```{r}
library(ggplot2)

# histogram
ggplot(data = mpg, mapping = aes(x = displ)) +
  geom_histogram()

# density plot
ggplot(data = mpg, mapping = aes(x = displ)) +
  geom_density()
```

# Two Variable Plots

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point()

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(color = "blue")

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(shape = 3)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(shape = 3, color = "blue")
```

# Grouped Plots

```{r}
# mean highway mpg by vehicle class
plot.data <- 
  mpg %>%
  group_by(class) %>%
  summarise(mpg = mean(hwy))
plot.data

ggplot(data = plot.data, mapping = aes(x = class, y = mpg)) +
  geom_col()

# distribution of highway mpg by vehicle class
ggplot(data = mpg, mapping = aes(x = hwy, fill = class)) +
  geom_histogram()

## this is kinda cluttered... let's dodge the columns
ggplot(data = mpg, mapping = aes(x = hwy, fill = class)) +
  geom_histogram(position = position_dodge())

# we also try faceting
ggplot(data = mpg, mapping = aes(x = hwy)) +
  geom_histogram() +
  facet_wrap(~ class)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) +
  geom_point()

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  facet_wrap(~ class)
```

# Customize Plot Appearance

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) +
  geom_point() +
  labs(x = "Engine Displacement", y = "Highway MPG", color = "Vehicle Class",
       title = "Highway MPG vs. Engine Size") +
  scale_color_discrete(labels = c("2-Seater", "Compact", "Midsize", "Minivan",
                                  "Pickup", "Subcompact", "SUV")) +
  theme_classic()
```


# Homework Problems

1. Create a histogram of cylinders. Why are there only three main spikes?


2. Create a scatter plot of city mpg and engine displacement.


3. Create a grouped bar plot of city mpg and vehicle class using the median 
instead of the mean. 


4. Create a scatter plot of city mpg and engine displacement and color the 
points red. Add labels to all of the axes. 


5. Create a scatter plot of city mpg and engine displacement and color the 
points by the vehicle's manufacturer. 