---
title: "Modeling in R"
output: html_notebook
---

# Set Working Directory

```{r}
setwd("/Users/nick/Documents/Teaching/workshops/data_science_series/5b_r_modeling")
```


# Import Data

```{r}
mpg <- read.csv("mpg.csv")
```


# Does the size of a car's engine affect highway MPG?

```{r}
model.1 <- lm(hwy ~ displ, data = mpg)
summary(model.1)
```


# Create a Table of Results

```{r}
library(stargazer)
stargazer(model.1, type = "text")

# let's add in variable names
stargazer(model.1, type = "text", covariate.labels = c("Engine Displacement"),
          dep.var.labels = c("Highway MPG"))
```


# Model with Controls

```{r}
model.2 <- lm(hwy ~ displ + year + cyl, data = mpg)
summary(model.2)

# Let's add this second model to our table
stargazer(model.1, model.2, type = "text", 
          covariate.labels = c("Engine Displacement", "Vehicle Year", 
                               "Number of Cylinders"), 
          dep.var.labels = c("Highway MPG"))

# To save the table to a word file, we use the 'out' argument
stargazer(model.1, model.2, type = "text", 
          covariate.labels = c("Engine Displacement", "Vehicle Year", 
                               "Number of Cylinders"), 
          dep.var.labels = c("Highway MPG"),
          out = "results_table.txt")
```


# Model Prediction

Let's use our model to predict the highway MPG for a vehicle with specific characteristics.

```{r}
# we start by creating a new dataframe with the characteristics of the car
# that we want to predict
pred.data <- data.frame(displ = 7, # car has 7 liters of engine displacement
                        year = 2008, # built in 2008
                        cyl = 8) # has 8 cylinders

predictions <- predict(object = model.2, newdata = pred.data, evel = 0.95, 
                       interval = "confidence")
predictions
```



# Beyond Linear Regression

```{r}
# Logit
fit <- glm(y ~ x1 + x2 + x3,
           family = binomial(link = "logit"),
           data = data)

# probit
fit <- glm(y ~ x1 + x2 + x3,
           family = binomial(link = "probit"),
           data = data)

# poisson
fit <- glm(y ~ x1 + x2 + x3,
           family = poisson(link = "log"),
           data = data)
```


# Homework Problems

1. Estimate a regression looking at the relationship between city MPG and 
engine displacement.


2. Estimate another model with year and the number of cylinders as controls.


3. Create a results table with both models and everything informatively labeled. 


4. Use your model to predict the city MPG for a car built in 1999, with 4 cylinders, and 1.8 liters of engine displacement. 