---
title: "Welcome to R"
output:
  html_document:
    df_print: paged
---

# This is a heading

## This is a smaller header

### This is an even smaller header

#### ... you get it

* This makes a bullet point

* What about a numbered list?

1. Just use numbers...

Here is how you create a block of R code:

```{r}

```

# Ok, Now Let's Code

R is just a really fancy calculator. Let's start by doing some basic calculations:

```{r}
# use "#" to write comments in a code block
# let's do some addition
1 + 1

5 + 2

# notice the spaces between the + operator. This is good grammar. It's a lot 
# harder to read without spaces (it gets worse when we have complex code).
1+1 # bad
5+2 # bad
read_excel(path="excel_lab.xlsx",sheet="cd_data") # bad

read_excel(path = "excel_lab.xlsx", sheet = "cd_data") # good

# multiplication
4 * 8
5 * 19

# division
4 / 8
14 / 7

# cubes
4 ^ 2
```

R shows the answers below the code block, but what if we wanted to save the answer to retrieve later? To do this, we can create an object but thinking of a name and using the assignment operator `<-`

```{r}
answer <- 5 + 2

# to see the answer we need to use the "print" function
print(answer)
```

If you look in the environment pane, you will see the object (answer) that we created.

# Combinations

What if we wanted to combine several operations in one object? We do this with the `c` function:

```{r}
# separate each operation with a comma
c(1 + 4, 4 * 2, 10)

# we can save this too
new.object <- c(1 + 4, 4 * 2, 10)
print(new.object)

# we can include text instead of numbers and operations. Put text in quotations
text <- c("This", "Shows", "That we can also have text")
print(text)
```

We can use this create variables.

# Types of Variables

In R, there are several different types of variables:

```{r}
# Numeric Variables
numbers <- c(1, 2, 3, 4, 5)
numbers
class(numbers)

# Character Variables
education <- c("Highshool", "Bachelors", "Grad School")
education
class(education)

# Logical Variables
male <- c(0, 1, 0, 1)
class(male)

male <- as.logical(male)
male
class(male)

# Categorical Variables
education <- as.factor(education)
```


# Functions

To see what arguments any function takes in R you can type a `?` before any function's name

```{r}
?mean()
```

# Data Exploration

## Working Directory

Now that we know a little bit about how R works, let's replicate everything we did in Stata. The first thing we did in Stata was set the working directory. To do this in R, we use the `setwd` function. 

```{r}
# what directory are we in now?
getwd()

# set the working directory
setwd(dir = "/Users/nick/Documents/Teaching/workshops/data_science_series/2b_r_intro")
```

## Import Data

To import our excel file with the data, we'll need to install and load the `readxl` package. This will give us the function we need to import excel files.

```{r}
# install (we only need to do this once)
install.packages("readxl")

# now that it's installed, we need to load it if we want to use it
library(readxl)

# now let's import the data set
read_excel(path = "excel_lab.xlsx", sheet = "cd_data")

# notice that I set the path and sheet arguments equal to something. We could
# have left these out, but then all the arguments must be entered in a specific
# order and it can be confusing to remember what each argument is doing.
read_excel("excel_lab.xlsx", "cd_data")

# this loads the data, but we didn't save it! We need to assign it to an object
data <- read_excel(path = "excel_lab.xlsx", sheet = "cd_data")

# now all the data is saved in an object called "data"
```

## 1. What does each row of data describe?

```{r}
View(data)
```

## 2. How many rows of data are there?

```{r}
nrow(data)
```

## 3. What is the average over 65 population across all congressional districts?

* To find the mean we use the function `mean`, but we also need to tell R what variable we want to find the mean for. The variable that we want is called `AGE65` and it is stored in the object we named `data`. First we give R the object, then we use `$` to find the variable. 

```{r}
mean(data$AGE65)
```

## 4. Create an indicator variable for California called "ca_dum"

We can also use `$` to add new variables:

```{r}
data$ca_dum <- ifelse(test = data$STATE == "CA", yes = 1, no = 0)
```

## 5. What is the average over 65 population in each state?

```{r}
aggregate(formula = FIPSTATE ~ AGE65, data = data, FUN = mean)
```

## 6. How many total blue collar workers are there in each state?

```{r}
aggregate(formula = FIPSTATE ~ BLUCLLR, data = data, FUN = sum)
```

# Homework Problems

## 1. Import the data on the "cd_data" sheet in the "excel_lab.xlsx" file


## 2. What is the average military population across all congressional districts.


## 3. Create an indicator called "cd_dum" for when a congressional district equals 10. 


## 4. How many total farmers are there in each state? (Name your variable "tot_farmers").