---
title: "Exploring Data with Excel"
author: MPP Workship Series
date: "October 1, 2020"
output: 
  ioslides_presentation:
    widescreen: TRUE
    incremental: TRUE
    theme: yeti
    logo: UCR_Icon_WhtBG.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Session Goals

In this session we'll work on the following:

* Practice identifying clean data

* Use excel to learn about using functions

* Use excel to answer several descriptive questions

## Identifying Clean Data

Review: 

1. We always want to figure out what each row tells us

2. We also want to figure out what each columns tells us

## Identifying Clean Data

```{r comment = ""}
head(ggplot2::mpg)
```

## Functions

Time for math! 

$$y = 5x$$

* If $x = 2$, then what does $y$ equal?

* We could also write this expression as $f(x) = 5x$

* $f(x)$ is a function. If we plug in a value for $x$ then we get a specific output

## Functions

* All data software programs use functions

* In excel, if we want to find the sum of numbers, we use the `SUM` function

  * `SUM(Number1)`
  
  * `AVERAGE(Number1)`
  
* All functions have inputs, or arguments

* In the $y = 5x$, $x$ is called the argument

* In the `SUM` and `AVERAGE` examples, `Number1` is the argument

## Time for Hands-on Practice!

Open `excel_lab.xlsx` on your computer and following along