---
title: "Data Science Workshop Assessment"
output: html_notebook
---

You've made it to the final assessment! In this final assignment you will need to build a model to predict the value of diamonds. This will require you to clean the data set and present your results. 

# Data Cleaning

1. Import the data

```{r}

```

2. Give all variables informative names. 

```{r}

```

3. Label the values of the `cut` variable.

```{r}

```

4. Make `cut`, `color`, and `clarity` categorical variables. 

```{r}

```


# Data Exploration

1. Get the summary statistics for all the variables in your data.

```{r}

```

2. Plot the distribution of diamond prices. 

```{r}

```

3. Plot the relationship between the clarity of a diamond and its price. 

```{r}

```


# Data Modeling

1. Build a model to predict the price of a diamond using the cut, carat, clarity, and color.

```{r}

```

2. Create an informative table of the model's results. 

```{r}

```

3. Create a coefficient plot of the model's results. 

```{r}

```

4. Predict the value of a 4 carat diamond with a very good cut, J color, and VVS2 clarity. 

```{r}

```