---
title: "Intro to Generalized Linear Models"
subtitle: "UCR GradQuant"
author: "Nicholas R. Jenkins"
date: "December 2, 2021"
output:
  ioslides_presentation: 
    widescreen: true
    logo: UCR_Icon_WhtBG.png
    incremental: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# The General Linear Model

## Forget generalizing it, what is the general linear model? {.build}

* The general linear model is just the same basic linear regression model you know and love. 

$$y = \alpha + \beta x_i + \epsilon$$

* But this way of writing regression equations hides a lot of assumptions

## Revealing the assumptions of your model {.build}

* To see what we're hiding, let's write this same model a different way

$$
\begin{aligned}
  y &\sim \text{Normal}(\mu_i, 0) \\
  \mu_i &= \alpha + \beta \times x_i
\end{aligned}
$$

* $\sim$ means "distributed"

* So, $y$ is normally distributed with a mean of $\mu$ and a standard deviation of 0

* We want to estimate the *average* outcome, which is why we "model" the mean

* To model the mean of our outcome, we use some linear equation: $\mu = \alpha + \beta * x$

## Revealing the assumptions of your model {.build}

So, 

$$
\begin{aligned}
  y &\sim \text{Normal}(\mu_i, 0) \\
  \mu_i &= \alpha + \beta \times x_i
\end{aligned}
$$

is the more complete way to express

$$y = \alpha + \beta x_i + \epsilon$$

because it tells us what likelihood and link (more on this later) functions we are using

## What are likelihood functions, again?

* Likelihood functions are essentially probability models of our data

* They describe the data generating process

* Ultimately the goal of statistics is to build a model that can make accurate predictions and describe data

* We do this by choosing likelihood functions that describe observed data

# Generalizing the General Linear Model

## What if the normal distribution doesn't describe our data?

```{r include = FALSE}
library(tidyverse)
library(nycflights13)
data(flights)
flights <- 
  flights %>% 
  mutate(flight_delayed = case_when(dep_delay > 0 ~ "Delayed",
                                    dep_delay < 0 ~ "Early"),
         delayed = ifelse(flight_delayed == "Delayed", 1, 0),
         distance_c = distance - mean(distance),
         distance_2sd = sjmisc::std(distance, robust = "2sd")) %>% 
  sample_n(size = 2000)
```

As an example, we're going to use data on flights in the US:

```{r}
head(flights)
```

## What if the normal distribution doesn't describe our data? {.build}

* Let's try to see if the distance of a flight can predict whether the flight is delayed or not

* What probability model would describe flight delays? The normal distribution?

## Modeling Flight Delays

```{r warning = FALSE}
ggplot(flights, aes(x = delayed)) +
  geom_density()
```

## Modeling Flight Delays {.build}

* The normal distribution doesn't sound right to me

* The normal distribution describes a data generating process that produces continuous values from $-\infty$ to $\infty$ 

* Let's run with it anyway

```{r echo = TRUE, eval = FALSE}
normal.fit <- glm(delayed ~ distance_2sd,
                  family = gaussian(),
                  data = flights)
```

## Modeling Flight Delays

```{r include = FALSE, warning = FALSE, message = FALSE}
library(brms)
library(cmdstanr)
normal.fit <- brm(delayed ~ distance_2sd,
                  family = gaussian(),
                  data = flights,
                  backend = "cmdstan",
                  cores = 4)
```

```{r warning = FALSE, message = FALSE}
library(bayesplot)
pp_check(normal.fit)
```

## Modeling Flight Delays {.build}

* The normal distribution doesn't model the data very well

* We need a likelihood function that describes a data generating process that produces discrete values of either 0 or 1

* The binomial and Bernoulli distributions both describe that process

## Modeling Flight Delays {.build}

$$
\begin{aligned}
  y_i &\sim \text{Binomial}(1, p_i) \\
  \text{logit}(p_i) &= \alpha + \beta \times x
\end{aligned}
$$

* What is our likelihood function? 

* What is our link function?

## Wait, what is a link function?

* Link functions are used to map the range of a linear equation to the range of the outcome variable

* The logit link function restricts the range of the linear equation to be either 0 or 1

* The log link function restricts the range of the linear equation to be positive, continuous, and greater than 0

## Connecting the Dots

* Our models always use the same linear structure ($\mu_i = \alpha + \beta x_i$)

* But, we can use any likelihood and link function that we want

* With these two pieces, we can **generalize** our linear equation to model any type of data

* Hence, the **generalized linear model**

* It always has a linear equation, but it doesn't always have a normal distribution!

## Modeling Flight Delays {.build}

* Let's model flight delays with a more realistic likelihood function - the binomial

$$
\begin{aligned}
  y_i &\sim \text{Binomial}(1, p_i) \\
  \text{logit}(p_i) &= \alpha + \beta \times x
\end{aligned}
$$

```{r echo = TRUE, eval = TRUE}
binomial.fit <- glm(delayed ~ distance_2sd,
                    family = binomial(link = "logit"),
                    data = flights)
```

## Modeling Flight Delays {.build}

```{r include = FALSE, warning = FALSE, message = FALSE}
binomial.bfit <- brm(delayed ~ distance_2sd,
                     family = bernoulli(link = "logit"),
                     data = flights,
                     backend = "cmdstan",
                     cores = 4)
```

```{r warning = FALSE, message = FALSE}
pp_check(binomial.bfit)
```

## How do we intrepret models with different likelihood and link functions?

* All you have to do is "undo" the link function

* Logit link function: $\ln \left( \frac{P}{1 - P} \right)$

* Inverse logit link function: $\frac{1}{1 + \exp^{(P)}}$

* Inverse logit link function in R: `plogis()`

## How do we intrepret models with different likelihood and link functions?

```{r}
coef(binomial.fit)
```

```{r include = TRUE, echo = TRUE}
plogis(coef(binomial.fit))
```

A one unit increase in `distance_2sd` increases the relative probability of having a delayed flight by about `r round(plogis(coef(binomial.fit)[[2]]), digits = 2) * 100`%

## The Flexibility of GLMs

* GLMs can be used to model any type of data (just about)

* Pick a realistic likelihood function, then pick a link function

## Modeling Test

```{r include = FALSE}
data("warpbreaks")

glimpse(warpbreaks)
```

We want to model yard breaks as a function of the type of wool and the tension applied. 

## Modeling Test

```{r}
hist(warpbreaks$breaks)
```

## Modeling Test

What kind of a distribution do we need?

* Continuous or discrete?

* Positive only, negative only, or all numbers?

## Modeling Test {.build}

* Poisson is discrete and positive only

* What kind of link function?

$$
\begin{aligned}
  y_i &\sim \text{Poisson}(\lambda_i) \\
  \text{log}(\lambda_i) &= \alpha + \beta \times x
\end{aligned}
$$

```{r echo = TRUE, eval = TRUE}
poisson.fit <- glm(breaks ~ wool + tension,
                   family = poisson(link = "log"),
                   data = warpbreaks)
```

## Modeling Test

What if we just used a normal likelihood function?

```{r include = FALSE, warning = FALSE, message = FALSE}
poisson.lin.bfit <- brm(breaks ~ wool + tension,
                        data = warpbreaks,
                        backend = "cmdstan",
                        cores = 4)
```

```{r warning = FALSE, message = FALSE}
pp_check(poisson.lin.bfit)
```

## Modeling Test

Now with a Poisson likelihood:

```{r include = FALSE, warning = FALSE, message = FALSE}
poisson.bfit <- brm(breaks ~ wool + tension,
                    family = negbinomial(link = "log"),
                    data = warpbreaks,
                    backend = "cmdstan",
                    cores = 4)
```

```{r warning = FALSE, message = FALSE}
pp_check(poisson.bfit)
```

## Modeling Test

How would we interpret the coefficients from a Poisson regression?

```{r}
exp(coef(poisson.fit))
```

* The coefficient was negative, so it decreases breaks

* Compared to wool A, the number of breaks for `woolB` is predicted to be $(1-0.813) * 100$ `r (1-0.813) * 100`% lower. 

# Let's practice!