---
title: "Cleaning in R"
output: html_notebook
---

# Overview

We're going to learn how to use R to complete all the cleaning tasks that we did in Stata.

# Set the Working Directory

```{r}
setwd("/Users/nick/Documents/Teaching/workshops/data_science_series/3b_r_cleaning")
```


# Importing the Data

Start by importing the data set. Don't forget that we need to load the `readxl` package to be able to import an excel file. 

```{r}
library(readxl)
leg.data <- read_excel("leg_data.xlsx")

# let's take a look
View(leg.data)

# Ok, same problems. We know what to do!
```

# You've heard about the Spiderverse, but what about Tidyverse?

Before we jump into it, we're going to learn about a style of syntax called "tidyverse." The tidyverse allows you to write R code in a way that makes it really easy to read. Let's see an example after we install and load the tidyverse package. 

```{r}
install.packages("tidyverse") # remember, we only need to install a package once
library(tidyverse)
```

Suppose we wanted to only look at the data for Republicans in our sample. Here are two ways we can do this:

```{r}
# without tidyverse
leg.data[leg.data$party == 1, ]

# with tidyverse
leg.data %>% filter(party == 1)
```

There are two pillars of the tidyverse. Using verbs to describe what our code is doing and the pipe operator (`%>%`). The five most common verbs are `filter`, `summarize`, `mutate`, and `arrange`. The pipe operator `%>%` can be read as "then." All that it does is let us go from one verb to the next. Here is how we would "say" the code below:

```{r}
# use "leg.data" THEN filter it for party equal to 1
leg.data %>% filter(party == 1)
```

We'll have more practice with this as we work through the Stata cleaning problems. 

# Rename Variables

Let's start by renaming some of our variables. 

```{r}
leg.data <- # save our changes to the object to leg.data
  leg.data %>% # use leg.data THEN
  rename(name = x, # rename the variable 'x' 'name'
         state = st,
         cong_dist = cd)

# take a look
view(leg.data)

# we can also look at the first few rows of the data a bit quicker with this:
head(leg.data)
```

# Value Labels

Now let's give the 1's and 2's of `party` more informative labels. To do this in R, we'll convert this variable to a categorical varible with the `factor` function. 

```{r}
leg.data <- # save our changes to the object to leg.data
  leg.data %>% # use leg.data THEN
  mutate(party = factor(x = party, # mutate creates new or changes existing variables
                        levels = c(1, 2), # these are the values in party
                        labels = c("Republican", "Democrat"))) # these are the labels we want to use

view(leg.data)

# notice that R automatically converted the value of -99 to "NA" for missing. 
# That's because we didn't list -99 as one of the category options. 
# Here is another way we could have fixed this missing value:
leg.data <- 
  leg.data %>%
  mutate(party = ifelse(test = party == -99, yes = NA, no = party))
# This says, "if the variable party equals -99 then change it to NA for missing
# but if it doesn't, then just use the value that it already is. 
```

# Using Other Tidyverse Verbs

## Summarize

Who's older in our data, Republicans or Democrats?

```{r}
leg.data %>% # no assignment here because we don't need to save our results
  group_by(party) %>% # group our data by the party variable
  summarize(M0974 = mean(M0974)) # calculate the mean of each group. 
```

## Arrange

What if we want to order our data by, say, age?

```{r}
leg.data %>% arrange(M0974)
```

# Bonus: Fixing State Abbreviations

Sometimes, we'd rather have full state names over abbreviations. In R there is a package called `openintro` that gives us a function to do just that. The function we need is `abbr2state`. 

```{r}
install.packages("openintro")
library(openintro)

leg.data <- 
  leg.data %>%
  mutate(state = abbr2state(abbr = state))
# fixed!
```



# Homework Problems

1. Give XMP098 and M094 more informative variable names.



2. Give the values "1" and "2" of XMP098 informative labels.



3. Fix the missing value of "-6" in XMP098 (You'll need to use `ifelse`)



4. What is the average birth year?



5. What is the average birth year by sex?